---
title: "Holt Winters"
author: "Lucia Ronchi Darre"
date: "11/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries

```{r}
library(ggplot2)
library(MASS)
library(fpp)
```

# Import data

Specify path:
```{r}
path <- "/Users/luciaronchi/Documents/MSCA/Time_Series/Final_Project/Data"
```

Load trips data:
```{r}
trips_data <- read.csv(paste(path, "RidershipTS.csv", sep = "/"), stringsAsFactors = FALSE)
```

# Explore and Prepare data

Cast Date variable as date type:
```{r}
trips_data$Date <- as.Date(trips_data$Date, "%m/%d/%Y")
```

Convert trips to time series for easier plotting:
```{r}
trips_ts <- ts(trips_data$number_of_trips, frequency = 365, start = c(2018, 305))
```

Plot trips over time:
```{r}
plot(trips_ts, main = "Daily Number of Rideshare Trips", xlab = "Time", ylab = "Number of Trips")
```

## Split into Train and Test

Select last day for the training period:
```{r}
train_end <- "2019-10-31"
```

Create train data:
```{r}
train_data <- trips_data[trips_data$Date <= train_end,]
```

Create test data:
```{r}
test_data <- trips_data[trips_data$Date > train_end,]
```

# Characteristics of the TS

* The TS seems to have a relatively stable level (no trend)
* The variance looks stable
* There is definitely daily seasonality


## Convert to TS object

Convert all data to weekly seasonal data:
```{r}
trips_ts <- ts(trips_data$number_of_trips, frequency = 7)
```

Convert train data to weekly seasonal data:
```{r}
train_ts <- ts(train_data$number_of_trips, frequency = 7)
```

Convert test data to weekly seasonal data:
```{r}
test_ts <- ts(test_data$number_of_trips, frequency = 7)
```

Test for stationarity:
```{r}
kpss.test(x = train_ts, null = "Trend")
```

* The null hypothesis of KPSS test is that the TS is stationary
* With a p-value of 0.01, the null hypothesis is rejected
* Thus, this TS needs to be made stationary

# Fitting an ADDITIVE HW on Train data

Fit an additive HW:
```{r}
m1 <- hw(y = train_ts, h = 61, exponential = FALSE, seasonal = "additive", damped = FALSE)
```

Model:
```{r}
m1$model
```

## RMSE in Train

Calculate RMSE
```{r}
sqrt(mean(m1$residuals**2))
```

## Plot forecasted Train data

Plot actual vs forecast:
```{r, fig.width = 3, fig.height = 1}
plot(train_ts, col = "blue", main = "Number of Trips Actual (Blue) vs Forecasted (Red)", ylab = "Number of Trips")
lines(m1$fitted, col = "red")
```

## Ljung Box Test for Residuals in Train

```{r}
Box.test(x = m1$residuals, type = "Ljung-Box", lag = 14)
```

## ACF of Residuals in Train

Plot ACF:
```{r}
acf(x = m1$residuals)
```

# Fitting a MULTIPLICATIVE HW on Train Data

Fit a multiplicative HW:
```{r}
m2 <- hw(y = train_ts, h = 61, exponential = FALSE, seasonal = "multiplicative", damped = FALSE)
```

Model:
```{r}
m2$model
```

## RMSE in Train

Calculate RMSE
```{r}
sqrt(mean(m2$residuals**2))
```

## Plot forecasted Train data

Plot actual vs forecast:
```{r, fig.width = 3, fig.height = 1}
plot(train_ts, col = "blue", main = "Number of Trips Actual (Blue) vs Forecasted (Red)", ylab = "Number of Trips")
lines(m2$fitted, col = "red")
```

We keep the ADDITIVE model as it yields a lower AICc
```{r}
m1$model$aicc < m2$model$aicc
```


# Predicting on Test data

Predict on test
```{r}
test_predictions <- predict(object = m1, new = test_ts)
```

Check number of predictions:
```{r}
length(test_predictions$mean)
```

## Plot Forecasted Test data

Transform to TS object:
```{r}
test_predictions_ts <- ts(test_predictions$mean, frequency = 7)
```

Plot predictions vs actual data:
```{r, fig.width = 3, fig.height = 1}
plot(test_ts, col = "blue", main = "Number of Trips: Actual (Blue) vs Forecast (Red)", ylab = "Number of Trips")
lines(test_predictions_ts, col = "red")
```

## RMSE and MAE in Test

Calculate RMSE in Test:
```{r}
sqrt(mean((test_predictions$mean - test_data$number_of_trips)**2))
```

MAE:
```{r}
mean(abs(test_predictions$mean - test_data$number_of_trips))
```

## ACF of Residuals in Test

Plot ACF:
```{r, fig.width = 3, fig.height = 1}
acf(x = (test_predictions$mean - test_data$number_of_trips))
```

## Ljung Box Test for Residuals in Test data

```{r}
Box.test(x = (test_predictions$mean - test_data$number_of_trips))
```




