---
title: "Time Series Final Project - Linear Regression"
author: "Lucia Ronchi Darre"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries

```{r}
library(ggplot2)
library(MASS)
library(fpp)
```

# Import data

Specify path:
```{r}
path <- "/Users/luciaronchi/Documents/MSCA/Time_Series/Final_Project/Data"
```

Load trips data:
```{r}
trips_data <- read.csv(paste(path, "RidershipTS.csv", sep = "/"), stringsAsFactors = FALSE)
```

Load Temperature, Precipitation, and Wind Speed data:
```{r}
weather_data <- read.csv(paste(path, "weather_final.csv", sep = "/"), stringsAsFactors = FALSE)
```

# Explore and Prepare data

## Temperature

Cast Date variable as date type:
```{r}
weather_data$DATE <- as.Date(weather_data$DATE, "%Y-%m-%d")
```

Plot histogram of temperatures:
```{r}
hist(weather_data$Daily.Avg.Temperature, main = "Distribution of Average Daily Temperature", ylab = "Frequency", xlab = "Temperature (F)")
```

## Precipitation

Plot histogram of precipitation:
```{r}
hist(weather_data$Daily.Total.Precipitation, main = "Distribution of Total Daily Temperature", ylab = "Frequency", xlab = "Precipitation (Inches)")
```

Given that the precipitation data is zero-inflated, we will convert it to binary:
```{r}
weather_data$precipitation_flag <- 0
weather_data$precipitation_flag[weather_data$Daily.Total.Precipitation > 0] <- 1
```

## Wind Speed

Plot histogram of Wind Speed:
```{r}
hist(weather_data$Daily.Avg.WindSpeed, main = "Distribution of Average Daily Wind Speed", ylab = "Frequency", xlab = "Wind Speed (km/h)")
```

## Trips

Cast Date variable as date type:
```{r}
trips_data$Date <- as.Date(trips_data$Date, "%m/%d/%Y")
```

Convert trips to time series for easier plotting:
```{r}
trips_ts <- ts(trips_data$number_of_trips, frequency = 365, start = c(2018, 305))
```

Plot trips over time:
```{r}
plot(trips_ts, main = "Daily Number of Rideshare Trips", xlab = "Time", ylab = "Number of Trips")
```

## Time

Create time-related variables that might be useful for the model:
```{r}
trips_data$day_of_week <- weekdays(as.Date(trips_data$Date))
trips_data$day_of_month <- format(trips_data$Date, format = "%d")
```

### Day of Week

Set days of the week as factors:
```{r}
trips_data <- within(trips_data, 
                   day_of_week <- factor(day_of_week, 
                                      levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))
```

Distribution for Day of Week:
```{r}
ggplot(trips_data[,c("number_of_trips", "day_of_week")], aes(x=day_of_week,y = number_of_trips))+ geom_bar(stat = "identity") + theme_bw() + 
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) 
```

### Day of Month

Distribution for Day of Week:
```{r}
ggplot(trips_data[,c("number_of_trips", "day_of_month")], aes(x=day_of_month,y = number_of_trips))+ geom_bar(stat = "identity") + theme_bw() + 
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) 
```

### End of Month

Create a variable that indicates that it's end of the month:
```{r}
trips_data$end_of_month <- 0
trips_data$end_of_month[trips_data$day_of_month >= 29] <- 1
```


### Weekend

Create a variable that groups Friday, Saturday, and Sunday:
```{r}
trips_data$weekend_flag <- 0
trips_data$weekend_flag[trips_data$day_of_week == "Friday" | trips_data$day_of_week == "Saturday" | trips_data$day_of_week == "Sunday" ] <- 1
```


# Merge and Prepare Data

## Merge data

Merge trips and weather data:
```{r}
data <- merge(x = trips_data, y = weather_data, by.x = "Date", by.y = "DATE")
```

## Remove unnecesary columns

Keep only certain columns:
```{r}
data <- data[c("number_of_trips", "Date", "day_of_week", "day_of_month", "end_of_month", "weekend_flag", "Daily.Avg.Temperature", "Daily.Avg.WindSpeed", "precipitation_flag")]
```

## Split into Train and Test

Select last day for the training period:
```{r}
train_end <- "2019-10-31"
```

Create train data:
```{r}
train_data <- data[data$Date <= train_end,]
```

Create test data:
```{r}
test_data <- data[data$Date > train_end,]
```


# Fit a Linear Regression on Train data

## Fit a full model

Fit linear regression with all variables as predictors:
```{r}
m1 <- lm(formula = "number_of_trips ~ .", data = train_data)
```

Get summary:
```{r}
summary(m1)
```

## Step AIC

Find the model with the best possible AIC:
```{r}
m2 <- stepAIC(object = m1, direction = "both", trace = TRUE, )
```

## Reduced model

Summary of the model found with stepAIC (it dropped day_of_month and WindSpeed)
```{r}
summary(m2)
```

## Plot Forecasted Train Data

Transform to TS object:
```{r}
train_predictions_ts <- ts(m2$fitted.values, frequency = 365, start = c(2018, 305))
train_ts <- ts(train_data$number_of_trips, frequency = 365, start = c(2018, 305))
```

Plot predictions vs actual data:
```{r, fig.width = 3, fig.height = 1}
plot(train_ts, col = "blue", main = "Number of Trips Actual (Blue) vs Forecasted (Red)", ylab = "Number of Trips")
lines(train_predictions_ts, col = "red")
```


## R2 in Train

```{r}
summary(m2)$adj.r.squared
```

## RMSE in Train

Calculate RMSE in Train:
```{r}
sqrt(sum((m2$residuals)**2))
```

## Ljung Box Test for Residuals in Train

```{r}
Box.test(x = m2$residuals, type = "Ljung-Box", lag = 10)
```

## ACF of Residuals in Train

```{r}
acf(x = m2$residuals)
```


# Forecast on Test Data

Forecast on test data using the best model found:
```{r}
test_predictions <- predict(object = m2, new = test_data)
```

## Plot Forecasted Test data

Transform to TS object:
```{r}
test_predictions_ts <- ts(test_predictions, frequency = 365, start = c(2019, 305))
```

Plot predictions vs actual data:
```{r}
plot(trips_ts, col = "blue", main = "Number of Trips: Actual (Blue) vs Forecast (Red)", ylab = "Number of Trips")
lines(test_predictions_ts, col = "red")
```

## RMSE in Test

Calculate RMSE in Test:
```{r}
sqrt(mean((test_predictions - test_data$number_of_trips)**2))
```



